---
title: "Examining Data from the Solar Photovoltaic Incentive Program beginning 2010"
author: "Jonathan Hernandez"
date: "August 1, 2016"
output: html_document
---

## The dataset that will be looked into is based on the Solar Photovoltaic 
## Incentive Program that has been completed among various Counties and Cities in
## New York State beginning in August 2010. This program is for installing grid-connected
## solar electric or photovoltaic systems for residential, commercial buildings/sectors.
## data that is included is the year is
##     - Project Installation Year
##     - PRoject Cost
##     - City and/or County location
##     - Expected Kwh Annual Production/Output


### Get the data from the website
```{r getDataSet, cache=TRUE,echo=TRUE}

# get URL of where the dataset is located and read it for analysis
URL <- "https://data.ny.gov/api/views/3pzs-2zsk/rows.csv?accessType=DOWNLOAD"
solar_PV <- read.csv(URL)
```

### Get some basic summary of the data like its structure and summary

```{r dataSummary, echo=TRUE}
# show the structure of the dataset
str(solar_PV)

# show some basic summary statistics on the data
summary(solar_PV)
```

## See the names and clean up the Column names if necessary
```{r cleanDataSet, echo=TRUE}
names(solar_PV)

# R takes characters like , and spaces and replaces them with .
# Lets rather replace those instead with _ character
colnames(solar_PV) <- gsub("\\.\\.*" , "_", colnames(solar_PV))

# to display the updated (cleaner) names
colnames(solar_PV)
```

### Examine some data with basic plots; show the log distribution of 
### Expected KWh Annual Production and boxplot based on year project was installed
### in homes/buildings

```{r simplePlots, echo=TRUE}

# show the (log) frequency histogram of Expected KWh Annual Production
# by taking the log of the Expected KWh Annual Production we can visualize
# better the data and see what kind of distribution it reveals
with(solar_PV,hist(log(Expected_KWh_Annual_Production)))

# now show a box plot of Expected_KWh_Annual_Production per year
# Color based on group for better reading and visulization
with(solar_PV,boxplot(log(Expected_KWh_Annual_Production) ~ Project_Install_Year,
                      xlab = "Project Install Year",
                      ylab = "Expected Kwh Annual Production",
                      col = c("red","blue","green","yellow","purple","orange"),
                      main = "Expected Kwh Annual Production grouped by Project Year"))
```

### show a scatterplot of the project count by city vs Project Cost

```{r scatterplot, echo=TRUE}
with(solar_PV,plot(Project_Count_by_City,Project_Cost,
                   xlab = "Project Count by City",
                   ylab = "Project Cost",
                   main = "Project Count by City vs Project Cost", pch = 8))

# draw a best fit line for the data so we can see if there is a strong relationship
abline(lm(solar_PV$Project_Cost ~ solar_PV$Project_Count_by_City), col = "red")
```

### Some things to note on these plots
####    - The Expected Kwh Annual Production is a right-skewed distribution
####    - Based on the box plot, the median Expected Annual Kwh Production was
####      about the same in 2011, 2012, 2014 and 2015. Also in 2014 and 2015,
####      Many projects were expected to produce much higher Kwh of solar energy
####      that in previous years
####    - The scatterplot of Project Count by City vs Project Cost shows a strong
####      positive linear relationship. Cities where there weren't many Projects
####      are heavily clustered around having cheap project costs.